---
title: "p8105 midterm project"
author: 'Zanis Fang, UID: ZF2213'
date: "10/16/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
 echo = TRUE,
 fig.width = 8,
 # fig proportion set to golden ratio
 fig.asp = 0.618,
 out.width = "100%"
)

# ggplot2 tibble dplyr tidyr ... etc.
library(tidyverse)

```


## Introduction

Accelerometers have been powerful tools to study personal activities over a long period of days. Unlike traditional ways of analysis, which focus mainly on the total activity over the day, analyzing activity over the course of the day can provide us more detailed information about the behaviours of the individuals, thus valuable healthcare information. This project aims at using both ways to analyse the accelerometer records of a congestive heart failure (CHF) patient, trying to wrangle valuable behaviour information about the patient.


## Data and Methods

* Patient information

The patient is a 63-year old male with BMI 25, who was diagnosed with congestive heart failure at Columbia University Medical Center. Consent from the patient is assumed since this information has already been public to whole class.

* Methods

The methods of this project are using simple r programming skills along with tidyverse packages to take a look at the activity distribution of days across whole period, within a day, and across days in a week. 


## Results

### Loading and tidying data

Columns are values rather than variables, so the first step to tidying is loading the data and gathering the columns into the minutes (epoch) of the days. Then rename some variables, make the variables with reasonable data type, arrange the columns and rows.

```{r loading_and_tidying, message = FALSE}
# loading data and make to tall table
activity <- read_csv("./data/p8105_mtp_data.csv") %>% 
	# gather into long table
	gather(key = "minute_in_day", value = "counts", activity.1:activity.1440) %>% 
	# get the "minute" of a day, 
	mutate(minute_in_day = str_replace(minute_in_day, "activity.", "")) %>% 
	# change to integer
	mutate(minute_in_day = as.integer(minute_in_day)) %>% 
	# make day in week variable factor and relevel the day variable
	rename(day_in_week = day) %>%
	mutate(day_in_week = forcats::fct_relevel(day_in_week, c("Sunday", "Monday",
																													 "Tuesday", "Wednesday",
																													 "Thursday", "Friday",
																													 "Saturday"))) %>% 
	# assuming the start of the first week is Sunday
	mutate(day = (week - 1) * 7 + as.integer(day_in_week)) %>% 
	# arrange the variable
	select(day, week, day_in_week, minute_in_day, counts)

```

After tidying, each row records the `r colnames(activity)` of an activity of a minute, there are in total `r nrow(activity)` minutes of `r max(activity$day)` days.

### The total activity in each day

```{r scatterplot_day_total}
activity_day_total <- activity %>% 
# aggregate across minutes to create total activity
	group_by(day, day_in_week, week) %>% 
	# get total activity in the days
  summarize(day_total = sum(counts))

# necessary package to put two ggplot elements together
library(patchwork)

# density plot of distribution of total activity across days
density_total <- activity_day_total %>% 
	# label the outliers
  group_by(outliers = day_total > 100000) %>% 
	# plot the density plot
	ggplot(aes(x = day_total)) +
	  geom_density(fill = "white") +
	  # draw a line to label the cutoff
	  geom_vline(xintercept = 100000, linetype = 2) +
	  # make coordinator flip
	  coord_flip() +
	  # label axis
	  labs(
	  	x = "Day total (counts)",
	  	y = "Density"
	  ) +
	  # ticks labels (before flip)
	  scale_x_continuous(labels = scales::comma) +
	  scale_y_continuous(breaks = c()) 

# day total counts over time
scatter_total <- activity_day_total %>%
	# group by outliers
	group_by(abnormal_day = day_total < 100000) %>%
	# plot scatterplot
	ggplot(aes(x = day, y = day_total, color = abnormal_day)) +
	  # scatterplot
	  geom_point(size = 2) +
	  # draw a line for cutoff
	  geom_hline(yintercept = 100000, linetype = 2) +
	  # get regression line for the days (exclude outliers)
	  geom_smooth(method = "lm", aes(linetype = abnormal_day), color = "black", show.legend = FALSE, se = FALSE) +
	  # labels, ticks, axis, titles, etc...
	  scale_linetype_manual(values = c(1, 0)) +
	  labs(x = "Day", color = "Abnormal Days", fill = "Abnormal Day",
	  		 title = "Figure.1 Total daily activity over time") +
	  theme(
	   axis.ticks.y = element_blank(),
	   axis.title.y = element_blank(),
	   axis.text.y = element_blank()
	  )

# put two plots together
density_total + scatter_total + plot_layout(widths = c(0.2, 1.8))
```

It turns out that in some days the total activity are extremely low, they are actually missing days (refer to supplementary material) of this time series observations. We can visually feel the daily total activity are increasing over time. For statistical analysis, it would be better removing NAs.

### Linear regression

Null hypothesis: slope is 0.
Alternative hypothesis: slope is not equal to 0.
Significance level: 0.01.

```{r linear_regression}
# get linear regression analysis
(stats_day_total <- summary(lm(day ~ day_total,
		 data = activity_day_total %>% filter(day_total > 100000))))

```

The p-value of correlation of total activities over time is `r format(stats_day_total$coefficients[2, 4], scientific = FALSE)`. So at 0.01 significance level, we have enough evidence to reject the null hypothesis and conclude that the patient became more active over time.

### The effect of days in a week on total activities.

```{r scatterplot_days_of_week}
# across week, show the distribution of activities for each day in a week
activity_day_total %>%
  # remove outliers
	filter(day_total > 100000) %>% 
	# boxplot of days in week
	ggplot(aes(y = day_total, x = day, group = day_in_week)) +
	  # boxplot
	  geom_boxplot() +
	  # scatterplot
	  geom_point(aes(color = day_in_week)) +
	  # linear regression line
	  geom_smooth(method = lm) +
	  # activity for each day in week
	  facet_grid(. ~ day_in_week) +
    scale_y_continuous(labels = scales::comma) +
	  labs(
	  	x = "Day",
	  	y = "Total activity in the days",
	  	title = "Figure.2 Activity of days in the week",
	  	color = "Day in week"
	  )

```

The increases of active during Friday, Saturday and Sunday over time are more significant that that of other days in the week. But the trends are all the same.

### Activities over the course of the day

```{r heatmap}
# display 24-hour profile for each day using 2d density plot
activity %>%
	# using heatmap to show every minute of each day
	ggplot(aes(y = minute_in_day / 60, x = day, color = log(counts, 2))) +
	  # heatmap
	  geom_point(size = 0.1, alpha = 0.5) +
	  # label special days
	  scale_x_continuous(breaks = c(1, 49, 55, 92, 108, 185, 243, 259, 300, 329)) +
		# hours in the days
	  scale_y_continuous(breaks = c(0, 8, 16, 24)) +
	  # palette
	  viridis::scale_color_viridis(
	  	name = "log2 Transformed Activity Intensity",
	  	option = "inferno",
	  	discrete = FALSE
	  ) +
	  # labels and themes
	  labs(
	  	y = "Time (hour)",
	  	x = "Day (with special days labelled)",
	  	title = "Figure.3 Heatmap of 24-hour activity profile for each day"
	  ) +
	  theme(
	  	axis.text.x = element_text(color = "red", angle = 45),
	  	legend.position = "bottom"
	  )

```

The patient are active during days and sedatory during night. He seems to live a regular life.

Regarding the data, there seem to be some "special" days (as mentioned in Figure.1), and some days with extremely low variance. Refer to supplementary material for more information.


### Daily distribution

```{r 24hour_day_in_week}
activity %>%
	# grouping to get number of a type of activity
	group_by(minute_in_day, day_in_week) %>%
	# get group total (adding the total activity of a single time point in a day across all days would be less sensitive to missing values)
	summarize(total_activity = sum(counts)) %>%
	# plot scatterplot and smoothline 
	ggplot(aes(x = minute_in_day / 60, y = total_activity, color = day_in_week)) +
	  # scatterplot
	  geom_point(size = 1, alpha = 0.6) +
	  # add smoothline
	  geom_smooth(color = "black", linetype = 1) +
	  # labels and axises
	  scale_x_continuous(breaks = c(0, 8, 16, 24)) +
	  scale_y_continuous(labels = scales::comma) +
	  labs(
	  	x = "Time (hour)",
	  	y = "Total Counts",
	  	color = "Day in Week",
	  	title = "Figure 4. 24-hour profile of different days in the week"
	  ) + 
	  # get the 24-hour profile for each day of the week
	  facet_grid(. ~ day_in_week)
```

The patient seems to be quite active during Friday and weekends. For some reasons, there are activity peaks during Thursday and Friday nights, and two peaks in Sundays.

## Discussion



## Supplementary material

```{r}
# some feature days
activity %>% 
	# get some days as examples
  filter(day %in% c(0, 49, 55, 92, 107:108, 185, 243, 244, 259, 300)) %>%
	# plot scatterplot for each days
	ggplot(aes(y = log10(counts), x = minute_in_day / 60, color = as.factor(day))) +
	  # scatterplot
	  geom_point(size = 0.5) +
	  # one plot for each day
	  facet_grid(day ~ .) +
	  # labels
	  labs(x = "Time (hour)",
	  		 y = "log10 Transformed Counts",
	  		 color = "Special Days",
	  		 title = "Supplementary figure.1 Some typical day observations")

```


